---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# netatmo.weather

<!-- badges: start -->
<!-- badges: end -->

The goal of netatmo.weather is to provide access to Netatmo measurements and station metadata making use of the Netatmo Weather API.

## Installation

You can install the development version of netatmo.weather with:

``` r
install.packages("devtools")
devtools::install_github("falk-env/netatmo.weather")
```

and load the package via

```{r}
library(netatmo.weather)
```

## Getting Started

### Authentication

In order to be able to access data, please follow the subsequent steps:

1) Register a user account at [auth.netatmo.com](https://auth.netatmo.com/access/signup) and login at [dev.netatmo.com](https://auth.netatmo.com/de-de/access/login?next_url=https%3A%2F%2Fdev.netatmo.com%2F).

2) Click on your username in the upper right corner and create a new application. Provide mandatory information (*) and save.

3) Credentials will be stored making use of `{keyring}`. To securely encrypt your secrets stored, it is necessary to define a master password in your R user environment, which can be edited via `file.edit("~/.Renviron")`. Create a new environment variable called `keyring_pw`. The new line added now looks something like `keyring_pw = "insert_your_strong_password_here_123!"`. You can also create a password using e.g. `{sodium}` with `keygen() |> bin2hex()`. Restart R to see changes.

4) Run `set_credentials()` and copy & paste the information necessary (name/id/secret), as supplied in 2) .

``` r
set_credentials()
#> Note: Keyring 'netatmo' successfully created.
```

5) Run `get_oauth2token()` to create an Oauth 2 token based on the specifications provided in 4).

```{r}
get_oauth2token()
```

6) When asked, whether you want to use a local file to cache OAuth access credentials between R sessions, choose "1: Yes".

7) You'll be redirected to your browser to grant access to your application. Accept and close the browser tab.

Note: `get_oauth2token()` is limited to `"read_station"` scope. Successful authentication is confirmed in the browser. Your token is now stored in `.httr-oauth` in your package directory and as `.sig` in your R environment. 

In case you wanted to execute **/getpublicdata** and **/getmeasure** API calls from your browser (for debugging reasons or whatever), you'll need to append your access token to your URL: `"&access_token=xxx"`. You'll also be notified if you try to execute requests with your access token missing.

You can access your access and refresh token consisting of a key and secret making use of little helpers provided:

``` r
print_at()
#> "&access_token=62361e03ca18e13802546z20|5dt2091f1693dbff35f0428f2386b492"

print_rt()
#> "&refresh_token=62361e03ca18e13802546z20|6ce2fb2490a615d58b16e874fz4eb579"
```

Token expire after 3 hours and have to be refreshed again in order to be used. `{netatmo.weather}` does so in the background automatically without the user noticing. But if you're using your access token in a browser session temporarily, make use of these little helpers provided:

``` r
is_expired()
#> TRUE

refresh_at()
#> <Token>
#> <oauth_endpoint>
#>  authorize: https://api.netatmo.net/oauth2/authorize
#>  access:    https://api.netatmo.net/oauth2/token
#> <oauth_app> de_uhi
#>   key:    62361e03ca18e13802546z20
#>   secret: <hidden>
#> <credentials> scope, access_token, expires_in, expire_in, refresh_token

is_expired()
#> FALSE
```

### /getpublicdata

Queries via `get_publicdata()` to obtain station locations and metadata require a user-defined bounding box as the main function argument. In order to facilitate this, `get_extent()` was implemented to help you out:

```{r}
# using coordinates (xmin, ymin, xmax, ymax)
e1 <- get_extent(x = c(6.89, 51.34, 7.13, 51.53), epsg = 4326)
e1

# using municipality names
e2 <- get_extent(x = "Essen")
e2

# using postal codes
e3 <- get_extent(x = "45145")
e3
```

This information can now be used to list stations located in this area (at the time of the query):

```{r}
stations <- get_publicdata(ext = e1)

# returning a simple feature collection with 309 features
dim(stations)

# inspect
head(stations, 10)
```

However, since the number of stations returned by /getpublicdata seems to be influenced by the size of the area queried, the logical argument `use_tiles` was implemented, slicing your area of interest in tiles à 0.05 degrees to be queried separately in order to ensure the maximum number of available stations.

```{r}
stations_tiled <- get_publicdata(ext = e1, 
                                 use_tiles = TRUE)

# returning a simple feature collection with 628 features using the same extent as above
dim(stations_tiled)

# inspect
head(stations_tiled, 10)
```

### /getmeasure

Queries via `get_measure()` to obtain station observations basically require a base station MAC address to be queried (included in `stations`), the parameter to be queried (e.g. `temperature`, `humudity`, `sum_rain`, `...`), the measurement interval in minutes (e.g. `5`, `30`, `60`) and a period encompassing the timestamp of the first and last observation to retrieve in form of the local UNIX time in seconds. 

To assist you with the latter going backwards from `Sys.time()`, `get_period()` exists:

```{r}
# per default returning the maximum number of observations (1024) as a function of `res` chosen
# default: `res = 5`
p1 <- get_period()
as.POSIXct(p1, origin = "1970-01-01")

# here: `res = 60` corresponding to hourly data
p2 <- get_period(res = 60)
as.POSIXct(p2, origin = "1970-01-01")

# querying the last 24 hours, maybe convenient for scheduled jobs
p3 <- get_period(x = "recent")
as.POSIXct(p3, origin = "1970-01-01")

# self-defined period
p4 <- get_period(x = c("2022-06-01", "2022-06-04"))
as.POSIXct(p4, origin = "1970-01-01")
```

This can now be used to acquire observations (iterating over previously identified stations) in form of listed `xts` objects. This might take some time to finish.

```{r}
# get subset of data for demonstration purposes
obs_temperature <- get_measure(devices = stations_tiled[1:10, ], 
                               period = p2, 
                               par = "temperature", 
                               res = 60)

class(obs_temperature)
length(obs_temperature)

# subset to individual xts object
xts <- obs_temperature[[1]]

class(xts)

# inspect index/coredata
head(xts)

# inspect attribute names appended 
attributes(xts) |> names() |> tail(-3)

plot(xts, main = "hourly air temperatures", col = "red")
```
